\section{Detailed Design}

\subsection{Significant algorithms}
\subsubsection{MonsterController}
\paragraph{Health and Life} Health determines how healthy a Monster is. It starts at 1 at birth and goes down to 0, at which point the monster dies. It will use the following formula:

Given \begin{math}d\end{math} is the date, \begin{math}b\end{math} is the monster's birthday, \begin{math}a\end{math} is the age rate, \begin{math}i\end{math} is the injury stat, \begin{math}B\end{math} is the base health and \begin{math}H\end{math} is the actual health:

\begin{math}
B = 2-exp((d-b)*a)\\
H = B-i
\end{math}

During battle, the amount of damage taken is decided from the other monster's strength stat, passed through a noise function that varies it.
\paragraph{Strength and Dodge} Strength is a stat that begins at 0 at birth, increases to max rate half-way through the age progression, then deteriorates again until reaching 0 at the end of the monster's natural life. The strength stat, combined with random noise, will determine how much damage will be dealt to the enemy monster per hit. It will use the following formula:

Given \begin{math}s\end{math} is the base strength and \begin{math}S\end{math} is the current strength:

\begin{math}
S = sB(exp(d-b)-1)
\end{math}

Dodge is a stat that, mathematically, is very similar to Strength. It determines the chances that a monster will dodge an attack. In order to determine whether the monster has dodged a given attack during battle, the dodge stat is combined with random noise and tested for whether it is above a given value. If it is, the dodge succeeds; if not, the damage is taken.
\paragraph{Breeding} Breeding will first test whether the monster will randomly mutate a stat and, if so, simply select a random value for the stat. If the monster will not randomly mutate the stat, it will select one of the two parents' stats to prefer, generate some noise for the stat and set that as the base of the stat for the new monster. The number of children generated will use the following formula:

given that \begin{math}n\end{math} is the number of children generated, \begin{math}M\end{math} is the maximum number of children any monster can have, \begin{math}f_{1}\end{math} is the fertility of the first parent and \begin{math}f_{2}\end{math} is the fertility of the second parent:

\begin{math}
n = M\sqrt{f_{1} * f_{2}}
\end{math}
\paragraph{Noise function} The noise function f will generate a random value between -1 and 1. It will be used in the following ways:
\begin{itemize}
\item{In a fight, the Strength attribute will be varied by \begin{math}S+(f*0.1)\end{math};}
\item{In a fight, the Dodge attribute will be varied by \begin{math}D+(f*0.5)\end{math};}
\item{In a fight, a successful Dodge will occur if the value returned by the above function is greater than 0.5;}
\item{During breeding, the preference of the parent for a given stat will be decided by whether the random number is positive or negative;}
\item{During breeding, a stat mutating randomly will be decided by whether the random number is less than -0.8;}
\item{During breeding, a mutating stat will be generated by \begin{math}\frac{f+1}{2}\end{math};}
\item{During the creation of a new random monster, the value of all stats will be decided by the above function.}
\end{itemize}
